\vfill A Survey of Web Testing Frameworks and a Workflow for Testing Javascript Applications

### 2014-05-15 Paper/paper.tex
More stuff.

    expect(12 - 2).toEqual(theAnswer);
  });
});
\end{lstlisting}
The ``describe'' statement creates a test suite to collect related tests under. Each ``it'' statement is generally one test with some number of ``expect'' statements. It may not be particularly pretty, but the above code is quite readable even to someone with no coding experience.

@@

Asynchronous testing using Jasmine was a bit of a chore, but there is support built in. This support comes in the form of a three-set function progression. The first function, ``runs(function () \{ .... \})'', runs the asynchronous code. The second function, ``waitsFor(function () \{ ... \})'', polls a flag that is set by the first function when it completes. Finally, the third function, another ``runs(...)'', houses the assertions to determine if the code ran properly or not. This method requires a fairly substantial amount of typing as well as complicated function chains when more than one asynchronous call is required.

@@

Mocha, by default, also uses the ``describe -> it(\"should...\" '' pattern to create tests. Mocha offers a variety of other test interfaces for developer's to write their test code the way they want to in addition to a custom, nearly unstructured variety. A few of the testing interfaces are outlined here:
\begin{lstlisting}
// BDD (default)
describe('BDD Suite', function () {
  it('should be test 1', function () {
    ...
  });
});

@@

Mocha's approach to asynchronous testing is quite a bit more straightforward than Jasmine's. In the ``it(...)'' anonymous function, there is passed an additional parameter called ``done'' which is then called when the asynchronous call has finished. This required us to edit the code we were testing to make all the async calls accept an additional callback, but it was a very small price to pay for extended testability. An example asynchronous test would look like this:
\begin{lstlisting}
describe("Asynch Mocha Test", function () {
  it("look like this", function (done) {
    myAsyncCall(param1, param2, function () {
      // This callback expected to run when async returns

@@

\paragraph{Assertion Library of Choice: Chai.js \cite{Chaijs}}
For the assertion library we chose Chai.js.  It offers the most flexibility for writing assertions that we'd found. It provides access to an Expect API, a Should API, and a more general Assert API. The Expect API provides assertions in the form of ``expect(myVar).to.equal(10);'' or ``expect(myArr).to.have.length(10);''. The Should API has assertions of the form ``myVar.should.be.a(\'number\');'' or ``myArr.should.have.length(10);''. Finally, the Assert API has the more tried and true ``assert.equal(myVar, 10);'' and ``assert.lengthOf(myArr, 10);''. The various APIs combined with Mocha's flexibilty led to a nice developer experience that can appeal to a variety of different developers.

\subsubsection{QUnit \cite{QUnit}}
QUnit is an Assertion style framework developed by the jQuery Foundation, the makers of jQuery, an immensely popular framework for DOM manipulation and access. QUnit is a bit more developer-centered and not as ``feel-good'' as the other frameworks but has, so far, been the most easy to use and has the most thorough test reporting of any framework out of the box. Mocha and Jasmine show how many ``specs'', the ``it'' functions, have passed, which is great information, but QUnit takes it a step further and breaks down each spec into the asserts contained within:
\begin{Figure}
  \centering
  \includegraphics[width=\linewidth]{qunitrunner.png} 
  \captionof{figure}{Each individual spec is further broken down into the assertions within it.}
\end{Figure}
An interesting feature that QUnit has that we'd not seen elsewhere is the ability to specify how many assertions are expected to run in the test. This is incredibly useful for callbacks. By setting an expected number of assertions, tests can no longer fail silently just because an assertion was never run. If the QUnit test finishes and the assertion count is not what was specified, it will fail the test and let the developer know fewer assertions were called than expected.

@@

QUnit's handling of asynchronous tests is similar to Mocha's except instead of adding a parameter to the ``it'' function, the developer calls the ``asyncTest'' function instead. Finally, the developer calls ``start()'' when the asynchronous test is finished. Example:
\begin{lstlisting}
asyncTest("Async Test", function () {
  myAsyncCall(param1, param2, function () {
    // This callback expected to run when async returns

@@

%-AMD
Intern uses Asynchronous Module Definition (AMD) form which is something we'd not seen before. Where this comes into play is merely in what used to be ``require(...)''. Instead of seperately requiring modules, the entire program is wrapped in a ``define([...])'' block that passes modules as parameters to an overarching function. That will make more sense after an example:
\begin{lstlisting}
// Intern
define([
    'intern!object',
    'intern/chai!assert',
    'app/hello'
], function (registerSuite, assert, hello) {
  // Test code here
});

@@

Getting Istanbul to output code coverage is as easy as installing it and then calling the ``istanbul'' function from the commandline with your test executable.
For example: ``istanbul cover mocha -- . -R spec'' will run the mocha tests in the current directory and output coverage statistics. To break that down a little further:

``istanbul cover'' tells istanbul to run coverage on the tests.

``mocha --'' specifies that the executable to be run is mocha and that the developer is done giving commandline args to istanbul, that's the ``--''

Finally, ``. -R spec'' is just the flags for the mocha command to run the appropriate test suites.

@@

\subsubsection{Grunt.js\cite{GruntJS}}
Grunt.js is a task-runner for Javascript. What Grunt does is automate routine tasks like linting, testing, copying files, and updating modules. The developer writes Grunt scripts that outline the necessary tasks and then can run them on the commandline with the ``grunt'' command.

Grunt is widely supported and has all kinds of built-in or installable plugins that make writing common tasks nearly trivial. As an example, a few built-in grunt tasks that we use include: ``jshint'' for linting, ``uglify'' for code minimization, and ``watch'' for automatically re-running tests when certain files change. Each of those plugins requires very little configuration and they're ready to be used.

Of particular interest is that last feature, ``watch''. We are particularly fond of grunt's ``watch'' capability because it means that we have tests running near constantly that give immediate feedback to the developer when they change existing code or write new code after having written a test for it. As long as all the tests are passing, the developer has a good idea that the changes made are regression-tested as well as freshly tested and are ready for publishing or integration.

\subsection{Misc Tools}
\subsubsection{Node Inspector\cite{NodeInspector}}
Node Inspector is a GitHub project for debugging Node.js server-side code. The inspector runs in the browser and is started by running ``node-debug'' instead of ``node'' when starting the application. Simple as that. Then you can set breakpoints using the normal browser developer tools.

@@

We decided to install everything globally because it made writing our scripts easier, but it's not a requirement. Everything can be done with locally installed node modules, instead of using the command by itself the developer would just need to supply the full path to the executable, which is generally ``PATH\_TO\_REPO/node\_modules/MODULE\_NAME/bin/MODULE\_NAME'' or something similar.

\subsubsection{Setting up Mocha}
To install Mocha, type into the terminal: ``npm install -g mocha''. The ``-g'' means to install the module globally.

Next, to install Chai, type into the terminal: ``npm install -g chai''.

To get Chai running with Mocha, just ``require(...)'' the appropriate chai assertion method in the test and start writing. For example, to get the Assert library: ``var assert = require(\'chai\').assert;''

After the tests have been written, the developer can run them using the ``mocha'' command. There are a number of command-line arguments available to the mocha command which are well documented and which we won't talk about here. We will talk about the one that we used consistently, that being ``-R''. The ``-R'' arguments allows the developer to specify how the test results will be formatted. Mocha offers seventeen different reporters with more reporters available online and the ability to create custom reporters.

@@

Fill out the fields with the appropriate values for a given project and click ``Create Project''. Depending on what the setup is for your project, you'll see slightly varying web pages after you create your project. One thing in common, though, is the button at the bottom of the steps list that says ``Customize Build''. After following the instructions, click ``Customize Build'':
\begin{Figure}
  \centering
  \includegraphics[width=0.95\linewidth]{magnumCI_customize_build.png}
\end{Figure}
There are a number of configuration options and panes that can be accessed on the left side. Our focus was making sure the ``Build Configuration'' tab was properly set up. This involved selecing the correct branch, the correct Node.js runtime version, the correct database from the ``Resources'' area, and adding the appropriate commands in the build steps. In our case we wanted ``mongoDB'' and ``Xvfb'' for UI testing, and then ``npm install -g'' in the ``Dependencies Installation'' to make sure our modules are available when we need them. Finally, we run nightwatch using ``nightwatch -c nightwatch.json'' (the nightwatch flags will be outlined in the next section) in the ``Test suite commands'' box. All of that makes our build configuration page look like this:
\begin{Figure}
  \centering
  \includegraphics[width=0.95\linewidth]{magnumCI_customize_build_filled.png}
\end{Figure}
After that setup, our project is ready to go. Anytime we commit from then on, MagnumCI will run and notify us of successful and unsuccessful builds.

\subsubsection{Setting up Nightwatch}

\subsubsection{Setting up Istanbul}

\subsubsection{Setting up Grunt}

\subsection{Workflow}
% - Unit Testing on save
% - Note OS X 'ulimit' issue with Grunt-watch
% - CI used for running browser tests

@@

% -- Extension to all of PX?
% -- Extension to non-proprietary open-source project
% -- Headless browsers

>>>>

    expect(12 - 2).toEqual(theAnswer);
  });
});
\end{lstlisting}
The \lstinline{describe} statement creates a test suite to collect related tests under. Each \lstinline{it} statement is generally one test with some number of \lstinline{expect} statements. It may not be particularly pretty, but the above code is quite readable even to someone with no coding experience.

@@

Asynchronous testing using Jasmine was a bit of a chore, but there is support built in. This support comes in the form of a three-set function progression. The first function, \lstinline!runs(function () { .... })!, runs the asynchronous code. The second function, \lstinline!waitsFor(function () { ... })!, polls a flag that is set by the first function when it completes. Finally, the third function, another \lstinline{runs(...)}, houses the assertions to determine if the code ran properly or not. This method requires a fairly substantial amount of typing as well as complicated function chains when more than one asynchronous call is required.

@@

Mocha, by default, also uses the \lstinline{describe -> it(\"should...\"} pattern to create tests. Mocha offers a variety of other test interfaces for developer's to write their test code the way they want to in addition to a custom, nearly unstructured variety. A few of the testing interfaces are outlined here:
\begin{lstlisting}
// BDD (default)
describe('BDD Suite', function () {
  it('should be test 1', function () {
    ...
  });
});

@@

Mocha's approach to asynchronous testing is quite a bit more straightforward than Jasmine's. In the \lstinline{it(...)} anonymous function, there is passed an additional parameter called \lstinline{done} which is then called when the asynchronous call has finished. This required us to edit the code we were testing to make all the async calls accept an additional callback, but it was a very small price to pay for extended testability. An example asynchronous test would look like this:
\begin{lstlisting}
describe("Asynch Mocha Test", function () {
  it("look like this", function (done) {
    myAsyncCall(param1, param2, function () {
      // This callback expected to run when async returns

@@

\paragraph{Assertion Library of Choice: Chai.js \cite{Chaijs}}
For the assertion library we chose Chai.js.  It offers the most flexibility for writing assertions that we'd found. It provides access to an Expect API, a Should API, and a more general Assert API. The Expect API provides assertions in the form of \lstinline{expect(myVar).to.equal(10);} or \lstinline{expect(myArr).to.have.length(10);}. The Should API has assertions of the form \lstinline{myVar.should.be.a(\'number\');} or \lstinline{myArr.should.have.length(10);}. Finally, the Assert API has the more tried and true \lstinline{assert.equal(myVar, 10);} and \lstinline{assert.lengthOf(myArr, 10);}. The various APIs combined with Mocha's flexibilty led to a nice developer experience that can appeal to a variety of different developers.

\subsubsection{QUnit \cite{QUnit}}
QUnit is an Assertion style framework developed by the jQuery Foundation, the makers of jQuery, an immensely popular framework for DOM manipulation and access. QUnit is a bit more developer-centered and not as ``feel-good'' as the other frameworks but has, so far, been the most easy to use and has the most thorough test reporting of any framework out of the box. Mocha and Jasmine show how many ``specs'', the \lstinline{it} functions, have passed, which is great information, but QUnit takes it a step further and breaks down each spec into the asserts contained within:
\begin{Figure}
  \centering
  \includegraphics[width=\linewidth]{qunitrunner.png} 
  \captionof{figure}{Each individual spec is further broken down into the assertions within it.}
\end{Figure}
An interesting feature that QUnit has that we'd not seen elsewhere is the ability to specify how many assertions are expected to run in the test. This is incredibly useful for callbacks. By setting an expected number of assertions, tests can no longer fail silently just because an assertion was never run. If the QUnit test finishes and the assertion count is not what was specified, it will fail the test and let the developer know fewer assertions were called than expected.

@@

QUnit's handling of asynchronous tests is similar to Mocha's except instead of adding a parameter to the \lstinline{it} function, the developer calls the \lstinline{asyncTest} function instead. Finally, the developer calls \lstinline{start()} when the asynchronous test is finished. Example:
\begin{lstlisting}
asyncTest("Async Test", function () {
  myAsyncCall(param1, param2, function () {
    // This callback expected to run when async returns

@@

%-AMD
Intern uses Asynchronous Module Definition (AMD) form which is something we'd not seen before. Where this comes into play is merely in what used to be \lstinline{require(...)}. Instead of seperately requiring modules, the entire program is wrapped in a \lstinline{define([...])} block that passes modules as parameters to an overarching function. That will make more sense after an example:
\begin{lstlisting}
// Intern
define([
    'intern!object',
    'intern/chai!assert',
    'app/hello'
], function (registerSuite, assert, hello) {
  // Test code here
});

@@

Getting Istanbul to output code coverage is as easy as installing it and then calling the \lstinline{istanbul} function from the commandline with your test executable.
For example: \lstinline{istanbul cover mocha -- . -R spec} will run the mocha tests in the current directory and output coverage statistics. To break that down a little further:

\lstinline{istanbul cover} tells istanbul to run coverage on the tests.

\lstinline{mocha --} specifies that the executable to be run is mocha and that the developer is done giving commandline args to istanbul, that's the \lstinline{--}

Finally, \lstinline{. -R spec} is just the flags for the mocha command to run the appropriate test suites.

@@

\subsubsection{Grunt.js\cite{GruntJS}}
Grunt.js is a task-runner for Javascript. What Grunt does is automate routine tasks like linting, testing, copying files, and updating modules. The developer writes Grunt scripts that outline the necessary tasks and then can run them on the commandline with the \lstinline{grunt} command.

Grunt is widely supported and has all kinds of built-in or installable plugins that make writing common tasks nearly trivial. As an example, a few built-in grunt tasks that we use include: \lstinline{jshint} for linting, \lstinline{uglify} for code minimization, and \lstinline{watch} for automatically re-running tests when certain files change. Each of those plugins requires very little configuration and they're ready to be used.

Of particular interest is that last feature, \lstinline{watch}. We are particularly fond of grunt's \lstinline{watch} capability because it means that we have tests running near constantly that give immediate feedback to the developer when they change existing code or write new code after having written a test for it. As long as all the tests are passing, the developer has a good idea that the changes made are regression-tested as well as freshly tested and are ready for publishing or integration.

\subsection{Misc Tools}
\subsubsection{Node Inspector\cite{NodeInspector}}
Node Inspector is a GitHub project for debugging Node.js server-side code. The inspector runs in the browser and is started by running \lstinline{node-debug} instead of \lstinline{node} when starting the application. Simple as that. Then you can set breakpoints using the normal browser developer tools.

@@

We decided to install everything globally because it made writing our scripts easier, but it's not a requirement. Everything can be done with locally installed node modules, instead of using the command by itself the developer would just need to supply the full path to the executable, which is generally \lstinline{PATH\_TO\_REPO/node\_modules/MODULE\_NAME/bin/MODULE\_NAME} or something similar.

\subsubsection{Setting up Mocha}
To install Mocha, type into the terminal: \lstinline{npm install -g mocha}. The \lstinline{-g} means to install the module globally.

Next, to install Chai, type into the terminal: \lstinline{npm install -g chai}.

To get Chai running with Mocha, just \lstinline{require(...)} the appropriate chai assertion method in the test and start writing. For example, to get the Assert library: \lstinline{var assert = require(\'chai\').assert;}

After the tests have been written, the developer can run them using the \lstinline{mocha} command. There are a number of command-line arguments available to the mocha command which are well documented and which we won't talk about here. We will talk about the one that we used consistently, that being \lstinline{-R}. The \lstinline{-R} arguments allows the developer to specify how the test results will be formatted. Mocha offers seventeen different reporters with more reporters available online and the ability to create custom reporters.

@@

Fill out the fields with the appropriate values for a given project and click ``Create Project''. Depending on what the setup is for your project, you'll see slightly varying web pages after you create your project. One thing in common, though, is the button at the bottom of the steps list that says ``Customize Build''. After following the instructions, click ``Customize Build'':
\begin{Figure}
  \centering
  \includegraphics[width=0.95\linewidth]{magnumCI_customize_build.png}
\end{Figure}
There are a number of configuration options and panes that can be accessed on the left side. Our focus was making sure the ``Build Configuration'' tab was properly set up. This involved selecing the correct branch, the correct Node.js runtime version, the correct database from the ``Resources'' area, and adding the appropriate commands in the build steps. In our case we wanted ``mongoDB'' and ``Xvfb'' for UI testing, and then \lstinline{npm install -g} in the ``Dependencies Installation'' to make sure our modules are available when we need them. Finally, we run Nightwatch using \lstinline{nightwatch -c nightwatch.json} (the nightwatch flags will be outlined in the next section) in the ``Test suite commands'' box. All of that makes our build configuration page look like this:
\begin{Figure}
  \centering
  \includegraphics[width=0.95\linewidth]{magnumCI_customize_build_filled.png}
\end{Figure}
After that setup, our project is ready to go. Anytime we commit from then on, MagnumCI will run and notify us of successful and unsuccessful builds.

\subsubsection{Setting up Nightwatch}
Installing Nightwatch consists of just installing another npm module. \lstinline{npm install -g nightwatch}. In addition to Nightwatch, the developer needs Selenium. The easiest way is to just download the jar straight from the Google repository here: ``http://selenium-release.storage.googleapis.com/index.html''. Once the developer has the JAR and places it somewhere, Nightwatch is ready to be configured and run. Nightwatch configuration happens in a json file that is pointed to when the developer executes Nightwatch.

The configuration file has plenty of options to get Nightwatch to do exactly what the developer wants, some of the more important ones include the \lstinline{selenium} object that holds information about where the executable is located and what ports to run on, the \lstinline{src\_folders} attribute to specify a source directory, and the \lstinline{test\_settings} object that contains configurations for default web pages, differing environments, and some desired capabilities.

To run Nightwatch, navigate to the project folder and run \lstinline{nightwatch -c path\_to\_nightwatch\_config}. This will use the \lstinline{src\_folders} attribute to determine where to look for the tests, start the Selenium server, and run the tests. When debugging certain functions of even test cases, tests can be isolated with the \lstinline{-t} flag: \lstinline{nightwatch -c nightwatch.json -t folder1/test1.js}. Nightwatch offers a number of other command-line arguments that are available on their devloper guide. \cite{NightwatchJS}

\subsubsection{Setting up Istanbul}
As mentioned in the Istanbul section, getting it running is simple. Install Istanbul with npm: \lstinline{npm install -g istanbul} and then run the istanbul command along with any others needed. In our case, we ran \lstinline{istanbul cover mocha -- . -R spec} to run Istanbul on our Mocha unit tests.

\subsubsection{Setting up Grunt}
Getting Grunt set up just the way the developer wants it is likely the most involved, but there are a myriad of examples and sample scripts to help them through. Installation is the same as we've seen before, \lstinline{npm install -g grunt-cli}. After installation the developer will need to create a ``Gruntfile'' which is just a Javascript file called ``Gruntfile.js''. Inside the Gruntfile are all of the command names and procedures that the developer wants. As the documentation outlines: ``A Gruntfile is comprised of the following parts: 1) The \"wrapper\" function 2) Project and task configuration 3) Loading Grunt plugins and tasks 4) Custom tasks''.

Here is an example Gruntfile:
\begin{lstlisting}
// 1) Wrapper Function
module.exports = function(grunt) {

  // 2) Project configuration.
  grunt.initConfig({
    pkg: grunt.file.readJSON('package.json'),
    uglify: {
      options: {
        banner: '/*! <%= pkg.name %> <%= grunt.template.today("yyyy-mm-dd") %> */\n'
      },
      build: {
        src: 'src/<%= pkg.name %>.js',
        dest: 'build/<%= pkg.name %>.min.js'
      }
    }
  });

  // 3) Load the plugin that provides the "uglify" task.
  grunt.loadNpmTasks('grunt-contrib-uglify');

  // 4) Default task(s).
  grunt.registerTask('default', ['uglify']);

};
\end{lstlisting}

A more complicatied Gruntfile simply consists of more tasks. Taken one at a time, it's easy to pick apart a Gruntfile. Task names can refer to one or more commands, so the command \lstinline{grunt testMyStuff} may run one command, or may run five commands. The task names are the ones that are used as the first parameter of the \lstinline{registerTask} call.

\subsection{Workflow}
In addition to just the a list of tools and installation instructions, we'd like to propose a way to use those tools that is efficient and nearly autonomous. Our goal is to bring testing tools into use with as little resistance as possible. With a bit of effort at the outset, testing can become a crucial and enjoyable part of devlopement.

What we recommend is a unit test suite that runs on save. While this may seem excessive, it provides invaluable, instant feedback for regression assurance and new functionality. The way to set this up with the above tools is have a grunt task that watches for changes in certain files and runs a script when the files change. Here's what our Grunt task looks like:
\begin{lstlisting}
module.exports = function(grunt)
{
  grunt.initconfig({
    watch: {
      test: {
          options: {
              spawn: false
          },
          files: ['**/*.js'],
          tasks: ['mochaTest']
      }
    },
    mochaTest: {
      test: {
          options: {
              reporter: 'spec',
              clearRequireCache: true
          },
          src: ['../../../../NetBeans/PolyXpressAuthor/test/unit/functional/*.js']
      }
    }
  });

  var defaultTestSrc = grunt.config('mochaTest.test.src');
  grunt.event.on('watch', function(action, filepath) {
      grunt.config('mochaTest.test.src', defaultTestSrc);
      if (filepath.match('test/functional')) {
        grunt.config('mochaTest.test.src', filepath);
      }
  });

  grunt.loadNpmTasks('grunt-contrib-watch');
  grunt.loadNpmTasks('grunt-mocha-test');

  // Register the mocha test task
  grunt.registerTask('test', ['mochaTest', 'watch:test']);
}
\end{lstlisting}

Our gruntfile is more complex than just that, but those are the parts that make the test-on-save functionality work. As we said earlier, Grunt has a lot of plugins that are available. We use one here to make our lives easier, \lstinline{grunt-mocha-test}. It is installed, once again, through npm and gives us access to the now-built-in task \lstinline{mochaTest}. The run-on-save functionality comes from the \lstinline{watch} tasks which has \lstinline{mochaTest} in its \lstinline{tasks} array.

Once that setup is done, the developer just runs \lstinline{grunt --force test} once which will execute the tests and then watch for changes and continue testing on save. The \lstinline{--force} flag is used so that the grunt task continues after ``failure'' which would occur if any tests did not pass.

% - Unit Testing on save
% - Note OS X 'ulimit' issue with Grunt-watch
% - CI used for running browser tests

@@

% -- Extension to all of PX?
% -- Extension to non-proprietary open-source project
% -- Headless browsers
% -- Sauce Labs
